---
title: "Applied Regression Analysis: HW2"
author: |
  | NCCU, Department of Statistics
  | 統計碩一 \ 葉佐晨
  | 112354016
date: "`r Sys.Date()`"
output:
  pdf_document:
    extra_dependencies:
      ctex: UTF8
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Variance Decomposition Proportions

## Packages required

```{r message=FALSE}
library(car)
library(magrittr)
library(mctest)
library(matlib)
library(knitr)
library(generics)
```

## Import data

Data is from **TABLE 9.4 Unstandardized Regressor and Response Variables from Webster, Gunst, and Mason [1974].**

```{r}
data <- read.table("tb9.4 Webster, Gunst, and Mason.txt", 
                   header = TRUE, 
                   row.names = "Observation")
```

## Center regressors

```{r}
X <- as.matrix(data[-1])
X_cen <- scale(X, scale = FALSE)
```

## Find eigenvectors

Show that **TABLE 9.6 Eigenvectors for the Webster, Gunst, and Mason Data**.

```{r}
X_corxz <- scale(X, scale = FALSE) %>% 
  apply(2, function(x) x/sqrt(sum(x^2)))

tb_eigvt <- eigen(t(X_corxz) %*% X_corxz) %>% 
  .[["vectors"]] %>% 
  round(5) %>% 
  as.data.frame
colnames(tb_eigvt) <- paste0("t", 1:6)

kable(tb_eigvt, longtable = FALSE, 
      caption = "TABLE 9.6 Eigenvectors for the Webster, Gunst, and Mason Data")
```

## Eigenvalues and Variance Decomposition Proportion

Show that **TABLE 9.7 Variance Decomposition Proportions for the Webster, Gunst, and Mason[1974] Data**.\
There are two methods can produce table 9.7, first is using package `mctest`, which is writed to do eigenvalues and variance decomposition proportion, second is caculating mathematically by hand.

### method 1, with `mctest`

```{r}
md_A <- lm(data$y ~ 0 + X_cen)
md_B <- lm(data$y ~ X)
```

#### A. Regressors Centered

```{r comment=""}
structure(eigprop(md_A))
```

#### B. Regressors Not Centered

```{r comment=""}
structure(eigprop(md_B))
```

### method 2, without `mctest`

#### A. Regressors Centered

```{r}
eigval_A <- eigen(t(X_corxz) %*% X_corxz)$values %>% round(5)

D_A <- svd(X_corxz)$d
eta_A <- (max(D_A)/D_A) %>% 
  round(5)

vif_A <- data.frame(y = data$y, X_corxz) %>% 
  lm(y ~ ., data = .) %>% 
  vif
t_A <- svd(X_corxz)$v

vdp_A <- t(t_A^2 %*% diag(1/D_A)^2) %*% diag(1/vif_A) %>% round(4)
dimnames(vdp_A) <- list(NULL,colnames(X))

tb_A <- data.frame(Number = 1:6,
                   Eigenvalue = eigval_A,
                   `Condition Indices` = eta_A,
                   vdp_A)
kable(tb_A, longtable = FALSE, caption = "TABLE 9.7: A. Regressors Centered")
```

#### B. Regressors Not Centered

```{r warning=FALSE}
X_corxz0 <- cbind(x0=1, X) %>% 
  apply(2, function(x) x/sqrt(sum(x^2)))
eigval_B <- (eigen(t(X_corxz0) %*% X_corxz0)$values)%>% 
  round(5)

D_B <- svd(X_corxz0)$d
eta_B <- (max(D_B)/D_B) %>% round(5)

t_B <- svd(X_corxz0)$v

vif_B <- data.frame(y = data$y, cbind(x0=1, X)) %>% 
  lm(y ~ 0 + ., data = .) %>% 
  vif
t <- svd(X_corxz)$v
vdp_B <- t(t_B^2 %*% diag(1/D_B)^2) %*% diag(1/vif_B) %>% round(4)
dimnames(vdp_B) <- list(NULL, colnames(X_corxz0))

tb_B <- data.frame(Number = 1:7,
                   Eigenvalue = eigval_B,
                   `Condition Indices` = eta_B,
                   vdp_B)

kable(tb_B, longtable = FALSE, caption = "TABLE 9.7: B. Regressors Not Centered ")
```
